|

### 2023-08-17 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
a bit more organizing and drafting. Very slow progress on this one

# Goals of chapter

- Demonstrate whether snap-judgments are used and how effective they are for groups of people
- Demonstrate how snap-judgments may be used even for non-political objects and the consequences
- Add one potential mechanism for the puzzle of partisan sorting: attitude expression suggests it happens, but behaviorally it does not happen much.
- Test the boundary conditions for color being a useful partisan cue

@@

This chapter builds on these findings to examine whether these associations between the colors red and blue with the partisanship of another can convey information about groups of people and can guide attitudes and resulting behaviors towards interactions with these groups of people. Specifically, the chapter builds upon a growing set of work suggesting that the politicultural differences between Democrats and Republicans are shaping seemingly non-political attitudes and behaviors such as one's choice in which neighborhood to live in. I argue in this chapter that the prevalence of red and blue cars in the driveways of houses in a neighborhood shift people's perceptions of the partisan composition of the neighborhood, and as a result, will make people less willing to move to that neighborhood when they perceive they'd be in the political minority of that neighborhood.

@@

Though not inherently or necessarily political, our communities often act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties [@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

The literature on partisan sorting seems to conclude that people express a desire to sort into neighborhoods with partisans, but these attitudes do not turn into observable behaviors. The remainder of the chapter provides one explanation for this puzzle by demonstrating how the snap-judgment model may be at play. Much like the literature on the attitude expression of partisan sorting, the snap-judgment model heavily relies on the ways in which seemingly non-political information comes to be associated with one partisan group or another. The model also demonstrates how these snap-judgments are capable of having initial impacts on attitude expression but its effects weaken in light of explicit and countervailing information.

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the types of cars that we drive are often used and rather effective cues to one's partisanship [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car.

Recall the snap-judgment model as depicted in @fig-theory. To apply it to the scenario of partisan sorting, absent other information about the neighborhood, the prevalence of particular colors of the cars in a neighborhood may be information that we can use to make guesses about the partisan composition of a neighborhood. Just as the other chapters demonstrated, when politics is salient, the associations between the colors red and blue trigger snap-judgments that inform attitude expression as well as behavioral reactions, and these effects appear to exist even once we have been given more explicit, and thus more informative, information about political persuasions.

>>>>

@@

In this chapter, I try to accomplish a few goals. First, I want to expand on the previous chapters by examining whether snap-judgments are used to evaluate groups of people. Second, I want to expand on the previous chapters by examining whether the previous findings extend to non-political objects. That is, this chapter seeks to further push my testing of the boundary conditions of the snap-judgment model. The third goal is that this chapter might address an important, but still yet to be solved, puzzle with the literature of partisan sorting that I discuss with more detail in the next section.

To achieve these three goals, the chapter builds upon a growing set of work suggesting that the politicultural differences between Democrats and Republicans are shaping seemingly non-political attitudes and behaviors such as one's choice in which neighborhood to live in. I argue in this chapter that the prevalence of red and blue cars in the driveways of houses in a neighborhood shift people's perceptions of the partisan composition of the neighborhood, and as a result, will make people less willing to move to that neighborhood when they perceive they'd be in the political minority of that neighborhood. While these snap-judgments influence attitude expression toward reported willingness to move to a neighborhood, subsequent information such as job opportunities, performance of schools, average wages to cost of living ratio, and other information limits the efficacy of these initial judgments. This seems to provide one mechanism explaining why partisan sorting is not observationally as extreme as we might expect given extensions of the literature on affective polarization and experimental and self-reported survey evidence.

@@

Though not inherently or necessarily political, our communities often collectively act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties [@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

The literature on partisan sorting seems to conclude that people express a desire to sort into neighborhoods with partisans, but these attitudes do not turn into observable behaviors. The remainder of the chapter provides one explanation for this puzzle by demonstrating how the snap-judgment model may be at play. Much like the literature on the attitude expression of partisan sorting, the snap-judgment model heavily relies on the ways in which seemingly non-political information comes to be associated with one partisan group or another. The model also demonstrates how these snap-judgments are capable of having initial impacts on attitude expression, but its effects weaken in light of explicit and countervailing information.

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the modal car in a neighborhood conveys the partisan composition of the whole neighborhood [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car. For a reader who thinks "hypothesizing about car color choice and partisanship is a reach", I agree. As stated at the outset of the chapter, the goal of the chapter is to extend past applications of the snap-judgment model to explicitly partisan objects such as yard signs as we did in Chapter 2 and to really challenge the model by looking at non-political objects that many never associate with partisanship.

Recall the snap-judgment model as depicted in @fig-theory. Extending the literature on partisanship as a superordinate social identity and affective polarization, we should expect that the partisan composition of a neighborhood would matter to people. In examining yard signs, @makse_politics_2019 demonstrate that people dislike having neighbors that do not share the same partisan affiliation as them. The motivation this comes from a desire to be in the majority of those around you with the same political views [see @van_der_does_strategic_2022]. While this appears to be something important to people, information about the partisan composition of a neighborhood might be inaccessible to most. Therefore, we are motivated to rely on heuristics to help with this [@motyl_how_2020]. One useful and efficient heuristic to convey partisanship is the use of color as other chapters have demonstrated. If we are concerned about the partisanship of our neighbors, we may rely on information such as the color of our potential neighbors' cars to infer their partisanship.

To most of us, we likely do not trust this as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, these usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Absent explicit information about the partisan composition, we might be willing to accept that the information we are basing our inferences on as
